---
title: "HW-4"
output: html_document
date: "`r Sys.Date()`"
---

Part I: Logistic Regression
1. Estimate a logistic regression model using "buyer" as the dependent variable and the following as predictor variables:
* last
* total_ 
* gender 
* child 
* youth 
* cook 
* do_it 
* reference 
* art
* geog

Hint: To do this in R, first transform the buyer and gender variables into a 0/1 dummy variable using data.table syntax
Then run the logistic regression command.
Finally, ask R to create a new variable that contains the predicted probability of purchase for each consumer.

```{r}
require(data.table)
bbb <- fread("~/Library/Mobile Documents/com~apple~CloudDocs/School Work/Sem 2/BUAN 6337/BUAN_6337_Predictive_Analysis/HW/HW-4/BBB.csv", 
             na.strings = c("NA", ""), 
               sep = "auto",
               stringsAsFactors = FALSE,
               data.table = TRUE,)
summary(bbb)

```

Changing buyer and gender variables into a flag

```{r}

bbb[,gender := ifelse(bbb$gender == "M",1,0)]
bbb[,buyer := ifelse(bbb$buyer == "yes",1,0)]

```

Logistic regression

```{r}
fmla <- buyer ~ last + total_ + gender + child + youth + cook + do_it + reference + art + geog
logit <- glm(fmla, data = bbb,family= binomial(link="logit"))

```
Predicted probabilities 

```{r}
bbb[, prob := predict(logit,type = "response")]
head(bbb)
```

2. Summarize and interpret the results (so that a marketing manager can understand them). Which variables are statistically significant? Which seem to be economically important? Interpret the odds-ratios for each of the predictors.

```{r}

summary(logit)

```


Part II: Decile Analysis of Logistic Regression Results
1. Assign each customer to a decile based on his or her predicted probability of purchase with ‘bucket 1’ being the highest average purchase probability.

```{r}
library(dplyr)
bbb[,decile := ntile(1 - bbb$prob,10)]
bbb[,bucket := paste("Bucket ",as.character(decile))]

```

2. Create a bar chart plotting the average response rate by decile (as just defined above).

```{r}

# Calculate the average response rate for each decile
response_rates <- aggregate(book_/total_ ~ decile, bbb, mean)

# Create the bar chart
hist(bbb$prob)


```