---
title: "HW_4_HXD220000_MXB220061_KXV220007_MXS220057"
output: html_document
date: "`r Sys.Date()`"
---

Authors : Mankirat Singh Bhamra MXB220061 - Harikrishna Dev HXD220000 - Krishnan Venkatesan KXV220007 - Medha Priyanga Saravanan MXS220057

Part I: Logistic Regression
1. Estimate a logistic regression model using "buyer" as the dependent variable and the following as predictor variables:
* last
* total_ 
* gender 
* child 
* youth 
* cook 
* do_it 
* reference 
* art
* geog

Hint: To do this in R, first transform the buyer and gender variables into a 0/1 dummy variable using data.table syntax
Then run the logistic regression command.
Finally, ask R to create a new variable that contains the predicted probability of purchase for each consumer.

```{r}
require(data.table)
bbb <- fread("~/Library/Mobile Documents/com~apple~CloudDocs/School Work/Sem 2/BUAN 6337/BUAN_6337_Predictive_Analysis/HW/HW-4/BBB.csv", 
             na.strings = c("NA", ""), 
               sep = "auto",
               stringsAsFactors = FALSE,
               data.table = TRUE,)
summary(bbb)

```

Changing buyer and gender variables into a flag

```{r}

bbb[,gender := ifelse(bbb$gender == "M",1,0)]
bbb[,buyer := ifelse(bbb$buyer == "yes",1,0)]

```

Logistic regression

```{r}
fmla <- buyer ~ last + total_ + gender + child + youth + cook + do_it + reference + art + geog
logit <- glm(fmla, data = bbb,family= binomial(link="logit"))

```
Predicted probabilities 

```{r}
bbb[, prob := predict(logit,type = "response")]
head(bbb)
```

2. Summarize and interpret the results (so that a marketing manager can understand them). Which variables are statistically significant? Which seem to be economically important? Interpret the odds-ratios for each of the predictors.

```{r}

summary(logit)

```

When we look at the z values of the different terms, we can determine the impact of the each variable in the logistic model. The art catgeory has the highest propensity of buyers as the |z value_art| is the highest. This is succeeded by the last time the person bought a book i.e. a buyer who hasn't bought a book in a while is more likely to buy a book now. Total value has the least significance among all the factors, whichc implies buyers don't care much about the total cost of their purchase.

Part II: Decile Analysis of Logistic Regression Results
1. Assign each customer to a decile based on his or her predicted probability of purchase with ‘bucket 1’ being the highest average purchase probability.

```{r}
library(dplyr)
bbb[,decile := ntile(1 - bbb$prob,10)]
bbb[,bucket := paste("Bucket ",as.character(decile))]
bbb[,decile := factor(decile,levels = c(1:10))]

```

2. Create a bar chart plotting the average response rate by decile (as just defined above).

```{r}

response_rates=bbb[,mean(buyer),by=decile]

barplot(V1~decile,data=response_rates)

```

3. Generate a table showing the number of customers, the number of buyers of "The Art History of Florence,' and the response rate to the offer by decile for the random sample (i.e. the 50,000 customers) in the dataset.

```{r}

results <- bbb[, .(num_cust = .N, num_buy = sum(buyer), resp_rate = sum(buyer) / .N), by = decile]

head(results)

```

4. For the 50,000 customers in the dataset, generate a table showing the average values of the following variables by probability of purchase decile:
Total $ spent
Months since last purchase, and
Number of books purchased for each of the seven categories (i.e., children, youth, cookbooks, do-it-yourself, reference, art and geography).

```{r}

avg_bbb <- bbb[, .(avg_total = mean(total_), avg_last = mean(last), 
                   avg_child = mean(child), avg_youth = mean(youth),
                   avg_cook = mean(cook), avg_do_it = mean(do_it),
                   avg_reference = mean(reference), avg_art = mean(art),
                   avg_geog = mean(geog)), 
              by = decile]

avg_bbb

```

5. Summarize and interpret the decile analysis results. Are the patterns in the decile analysis consistent with your conclusions from the logistic regression? (Hint: graph some of the results in the previous question.)

```{r}
# summarize average values by decile
require(psych)
results2 <- describeBy(bbb,bbb$decile)
results2
```
```{r}

plot(avg_bbb$decile,avg_bbb$avg_total,main = "Purchase ammount",xlab = "Deciles",ylab = "Average price")
plot(avg_bbb$decile,avg_bbb$avg_last,main = "Last month",xlab = "Deciles",ylab = "Number of months since most recent purchase")
plot(avg_bbb$decile,avg_bbb$avg_child,main = "Children Book",xlab = "Deciles",ylab = "Average number of children's books purchased")
plot(avg_bbb$decile,avg_bbb$avg_youth,main = "Youth books",xlab = "Deciles",ylab = "Average number of youth books purchased")
plot(avg_bbb$decile,avg_bbb$avg_cook,main = "Cook books",xlab = "Deciles",ylab = "Average number of cook books purchased")
plot(avg_bbb$decile,avg_bbb$avg_do_it,main = "Do it books",xlab = "Deciles",ylab = "Average number of do it books purchased")
plot(avg_bbb$decile,avg_bbb$avg_reference,main = "Reference books purchased",xlab = "Deciles",ylab = "Average number of reference books purchased")
plot(avg_bbb$decile,avg_bbb$avg_art,main = "Art books",xlab = "Deciles",ylab = "Average number of art books purchased")

```
Total:
We can see that Top deciles purchase the most in terms of purchase amount and it decreases as the decile progress. Note that there is small rise in Decile 10 but that could be due to population bias.

Last month:
We can see that recency of purchase heavily influences probability of purchase as it more likely to purhcase the book if its been a long time since the customers last purchase.

Children category:
We see that its evenly distributed around decile 5, where customers are more likely to buy the book when they are closer to decile 1 or 10. 

Youth category:
We see that its almost evenly distributed around decile 5 but it peaks at Decile 1.

Cook category:
We see that its evenly distributed around decile 5, where customers are more likely to buy the book when they are closer to decile 10.

Do it category:
We can see that do it books are most purchased by Decile 10 and the probability of do it books being bought dips at decile 4.

Reference category:
We can see that do it books are most purchased by Decile 1 and the probability of do it books being bought dips at decile 8.

Art books:
People who buy the book have a higher probability of buying other art books.


Part III: Profitability Analysis 
Use the following cost information to assess the profitability of using logistic regression to 
determine which of the remaining 500,000 customers should receive a specific offer: 
Cost to mail offer to customer: 0.5$
Selling price (shipping included): 18$
Wholesale price paid by BookBinders: 9$
Shipping costs: 3$
1. What is the breakeven response rate? 

The breakeven response rate can be calculated as follows:
Revenue per customer = Selling price - Wholesale price - Shipping costs = $18.00 - $9.00 - $3.00 = $6.00
Breakeven cost per response = Cost to mail offer to customer / Revenue per customer = $0.50 / $6.00 = <b>0.0833</b>
Therefore, the breakeven response rate is <b>8.33%</b> (i.e., we need at least 8.33% of the targeted customers to respond to the offer to break even).

2. For the customers in the dataset, create a new variable (call it "mailto_logit") with a value of 1 
if the customer's predicted probability is greater than or equal to the breakeven response 
rate and 0 otherwise. 
Hint: You can use data.table’s conditional assignment syntax

```{r}
# calculate breakeven response rate
revenue_per_customer <- 18 - 9 - 3
breakeven_cost_per_response <- 0.5 / revenue_per_customer
breakeven_response_rate <- breakeven_cost_per_response * 100
breakeven_response_rate

# create new variable "mailto_logit"
bbb[, mailto_logit := ifelse(prob >= breakeven_cost_per_response, 1, 0)]
bbb
```

3. Out of the 50,000 test sample, how many customers should have received the targeting 
promotion mail for “The Art History of Florence” based on the breakeven response rate (i.e., 
the number of mailto_logit ==1 in the data)? Also, among those who would have targeted, 
what would have been the response rate (i.e., mean of buyer among mailto_logit ==1)? How 
much higher is this response rate relative to the overall response rate in the data?

```{r}
# calculate number of customers to receive offer
num_customers <- sum(bbb$mailto_logit)
num_customers

# calculate response rate among those who receive offer
response_rate <- mean(bbb$buyer[bbb$mailto_logit == 1])
response_rate

# calculate expected profit from campaign
expected_profit <- num_customers * (18 - 9 - 3 - 0.5)
expected_profit
```

4. Consider that there are 500,000 remaining customers for the roll-out (excluding 50,000 test 
group in the current data. Assuming our test group is similar to the roll-out group (i.e., our 
test group is representative of the roll-out group), what is the expected number of buyers of 
'The Art History of Florence' if we do targeted mailing based on the breakeven response rate?
Hint: Count the number of buyers among the targeted (mailto_logit==1) in the test group and 
multiply it by 10 since the roll-out sample is 10 times larger. Alternatively, you can use the 
response rate.

```{r}

# calculate expected number of buyers in roll-out group
expected_num_buyers <- num_customers * (mean(bbb$buyer[bbb$mailto_logit == 1]) / mean(bbb$mailto_logit))
expected_num_buyers_rollout <- expected_num_buyers * 10
expected_num_buyers_rollout

```


