---
title: "HW2_6337_2023"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
date: '`r Sys.Date()`'
editor_options: 
  markdown: 
    wrap: 72
---

Authors : Mankirat Singh Bhamra MXB220061 - Harikrishna Dev HXD220000 -
Sarthak Vajpayee SXV220020 - Krishnan Venkatesan KXV220007

Loading required libraries and cleaning environment

```{r}
rm(list = ls())
demo = T
require(psych)
require(data.table)
require(dplyr)
require(ggplot2)
require(rlist)
if(demo) {setwd("~/Library/Mobile Documents/com~apple~CloudDocs/School Work/Sem 2/BUAN 6337/BUAN_6337_Predictive_Analysis/HW/HW-3")}
```

1 (R) (Weak) Law of Large Numbers • (a) Consider a continuous random
variable Xi ∼ Uni f orm [0, 2]. What is E [Xi] and Var(Xi)?

Ans: The mean (expected value) of a continuous uniform distribution [a, b] is
given by:

$$   E[X] = \frac{(a + b)}{2} $$

In this case, a = 0 and b = 2, so we have:

$$  E[Xi] = \frac{(0 + 2)} {2}= 1 $$

Therefore, the expected value of Xi is 1.

The variance of a continuous uniform distribution [a, b] is given by:

$$  Var[X] = \frac{(b - a)^2}{12} $$

In this case, a = 0 and b = 2, so we have:

$$  Var[Xi] = \frac{(2 - 0)^2}{12}= \frac{1}{3} $$

Therefore, the variance of Xi is $\frac{1}{3}$.

• (b) Consider the Xi defined above in (a) for i = 1, 2, ..., n, where
each Xi ⊥ Xj whenever i ̸= j. Consider the sample mean

What is E [X ̄ n] and Var (X ̄ n)? (The answer might be a function of
n).

Ans: The sample mean of n independent and identically distributed (i.i.d.) random variables Xi has the following properties:

$$ E[X̄n] = E[\frac{1}{n} ∑X_i] =\frac{1}{n} ∑E[X_i] =\frac{1}{n} \times n\times E[X_i] = E[X_i]$$

$$Var[X̄n] = Var[\frac{1} {n} ∑X_i] =\frac{1}{n^2} ∑Var[X_i]$$

Since each Xi is uniformly distributed on [0, 2], we know from part (a)
that $E[Xi] = 1$ and $Var[Xi] = \frac{1}{3}$. Therefore:

$$E[X̄n] = E[Xi] = 1$$

$$ Var[X̄n] = Var[\frac {1} {n} ∑X_i] =\frac{1}{n^2} ∑Var[X_i] = \frac{1}{n^2}\times n\times\frac{1} {3}= \frac{1} {3n}$$

Therefore, the expected value of the sample mean X̄n is 1, and the
variance of the sample mean X̄n is $\frac{1}{3n}$.

Now repeat the following (c)-(e) for
$n = 1, 2, 3, 5, 10, 50, 100, 1000, 3000$. Use the for loops to execute
(c)-(e). • (c) (R) Generate a size n vector of independent Uni f orm [0,
2] random variables and cal- culate its sample mean X ̄ n. • (d)(R)Take
$|X ̄n−E[Xi]|$ and report the value. • (e) (R) Now consider a continuous
transformation \$ f(x) = 2x\^2 − 5x + 1 + 1 \$. Take
$| f ( X ̄ n ) − f ( E [ X i ] ) |$

```{r}
n = c(1, 2, 3, 5, 10, 50, 100, 1000, 3000)

Q1 = function(n){
  
  # c
  x = runif(n,0,2)
  x_bar = mean(x)
  
  #d
  e_x = 1
  diff = abs(x_bar - e_x)
  
  #e
  quad = function(x)
  {
    return (2*x^2 - 5*x + 1 + (1/(3*x)))
  }
  
  diff_function = abs(quad(x_bar)-quad(e_x))
  return(list(n=n,mean=x_bar,diff=diff,diff_func=diff_function))
}

for(i in n){
  assign(paste0("v",i),Q1(i))
  if(demo){print(get(paste0("v",i)))}
  
}

```

• (f) What happens to the reported value in (d) and (e) as n increases?
Discuss.

Ans: As n increases, the difference between the theoretical value and practical value tends towards 0. This implies larger the sample, the results are better statistically.

```{r}

diff_func = c()
for(i in n){
  diff_func = append((Q1(i)$diff_func),diff_func)
}
plot(rev(diff_func))

```

2 (R) The Central Limit Theorem • (a) Consider a continuous random
variable $Xi ∼ Uniform [0, 2]$. What is $E[X_i]$ and $Var(X_i)$?

$$ Mean = \frac {a+b}{2} = 1 $$ $$ Var = \frac {1}{3} $$

• (b) Consider the Xi defined above in (a) for i = 1, 2, ..., n, where
each $Xi ⊥ Xj$ whenever $i \ne j$. Consider the sample mean ̄1n Xn := n
∑Xi. i=1 What is $E[X ̄ n]$ and $Var(X ̄ n)$? (The answer would be a
function of n).

Ans: 

$$ E[X_n] = \frac{1}{n} \times {n} \times 1 = 1 $$
$$ E[X_{n=6}] = 1 $$

$$ Var[X_n] = \frac{1}{\sqrt{n}} \times n \times 0.3333 = 0.041667 $$

• (c) Consider the transformation
$$ Y_n : = \sqrt n \times ( X _n − E[ X_i ] ) $$ What is E [Yn] and Var
(Yn)?

Ans:

$$E[Y_n] = \sqrt {n} \times 0 = 0  $$

$$ var = \frac {1}{3} $$

• (d) Consider the transformation √ ( X ̄ n − E [ X i ] ) Zn:=
npVar(Xi). What is E [Zn] and Var (Zn)?

$$ E[Z_n] = 0 $$

$$ Var[Z_n] = 1 $$

Now repeat the following for n = 1, 2, 3, 5, 10, 50, 100, 1000, 3000.
Use the for loops to execute (e)-(m). • (e)(R)Generate t =
1,2,3,4,5,6,7,8,9,10,11,...,2500 n×1 vectors of independent
$Uniform[0,2]$ random variables and calculate its sample mean $X ̄ nt$
respectively for each t. Denote this size 2500 vector as
$$v_2500 ≡X ̄_1,X ̄_2,...,X ̄_2500$$ for now.

```{r}
n = c(1, 2, 3, 5, 10, 50, 100, 1000, 3000)

Q2 = function(n){

  t = c(1:2500)
  v = replicate(length(t), runif(i, 0, 2))
  if (n == 1) {
    vn_2500 = v
  } else {
    vn_2500 = apply(v, 2, mean)
  }
  
  
  vn_2500_bar = mean(vn_2500)
  vn_2500_var = var(vn_2500)
  
  e_xi = 1
  yn_2500 = sqrt(n)*(vn_2500 - e_xi)
  
  mean_yn_2500 = mean(yn_2500)
  var_yn_2500 = var(yn_2500)
  
  var_xi = 1/3
  zn_2500 = sqrt(n)*(vn_2500 - e_xi)/sqrt(var_xi)
  
  mean_zn_2500 = mean(zn_2500)
  var_zn_2500 = var(zn_2500)
  
  return(list(n=n, vn = vn_2500, mean_vn = vn_2500_bar, var_vn = vn_2500_var, zn = zn_2500, mean_zn = mean_zn_2500,yn = yn_2500, var_zn = var_zn_2500))
}
for(i in n){
  assign(paste0("f",i),Q2(i))
}
for(i in n){
  hist(Q2(i)$zn, breaks = 100,
       main = paste("Histogram of zn_2500 for n = ", i, sep = ""),
       xlab = "zn_t values",
       xlim = c(-3,3))
}
for(i in n){
  hist(Q2(i)$yn, breaks = 100,
       main = paste("Histogram of zn_2500 for n = ", i, sep = ""),
       xlab = "vn_t values",
       xlim = c(-3,3))
}
```

• (m) What happens to the reported values in (f), (h), (k) and
histograms in (i) and (l) as n increases? Discuss.

3 (R) WLLN with Simple Regression

In this exercise, you will generate datasets for simple regression
yourself, and then try to esti- mate the model parmaters to examine the
properties of simple regression OLS estimators as the sample size n
grows. Repeat the following for n = 2, 3, 4, 5, 6, 7, 8, 9, 10, 15, 20,
30, 50, 75, 100, 250, 500, 1000, 2000, 3000. Use the for loops as
necessary.

• (a) (R) Generate a n × 1 column vector of Uniform[0, 12] random
variable and denote it as x.

• (b) (R) Generate a n × 1 column vector of Uni f orm [−4, 4] random
variable and denote it as u. • (c) (R) Generate the y vector using the
following formula: yi = 3 + 2xi + ui for each i = 1, 2, 3, ..., n. That
is, i'th row of x and u corresponds to i'th observation. • (d) (R) Now
you have a Monte-Carlo dataset of size n. Estimate the β in the
following model using OLS. (Recall the formula d .) What is the
calculated value of βOLS,n? Report.

```{r}

n = c(2, 3, 4, 5, 6, 7, 8, 9, 10, 15, 20, 30, 50, 75, 100, 250, 500, 1000, 2000, 3000)

Q3 <- function(n) {
  
  x = matrix(runif(n, 0, 12), nrow = i, ncol = 1)
  
  u = matrix(runif(n, -4, 4), nrow = i, ncol = 1)
  
  y = 3 + (2 * x) + u
  
  cov_xy = cov(x, y)
  var_x = var(x)
  hat_beta = cov_xy/var_x
  
  abs_beta = abs(hat_beta - 2)
  
  return(list(n=n,x=x,u=u,y=y,cov_xy=cov_xy,var_x=var_x,hat_beta=hat_beta,abs_beta=abs_beta))
}

for(i in n){
  assign(paste0("x",i),Q3(i))
}
```

4 In this exercise, you will generate datasets for simple regression
yourself, and then try to esti- mate the model parmaters to examine the
properties of simple regression OLS estimators as the sample size n
grows. Repeat the following for n = 2, 3, 4, 5, 6, 7, 8, 9, 10, 15, 20,
30, 50, 75, 100, 250, 500, 1000, 2000, 3000. Use the for loops as
necessary.

• (a) (R) Generate a n × 1 column vector of Uni f orm [0, 12] random
variable and denote it as x. • (b) (R) Generate a n × 1 column vector of
Uni f orm [−4, 4] random variable and denote it as u. • (c) (R) Generate
the y vector using the following formula: yi = 3 + 2xi + ui for each i =
1, 2, 3, ..., n. That is, i'th row of x and u corresponds to i'th
observation. • (d) (R) Now you have a Monte-Carlo dataset of size n.
Estimate the β in the following model $$yi = α + βxi + ui $$

• (e) (R) Repeat (a)-(d) for 2,500 times. You must have 2,500 βˆOLS,n
estimates in the memory ˆ1 ˆ2 ˆ2500 ′ at the end of this sub-question.
Denote this size 2,500 vector by b = βOLS,n, βOLS,n, ..., βOLS,n . • (f)
(R) Calculate the variance of b and report. • (g) (R) Subtract 2 from b
and multiply (c1, c2, ...., c2500), i.e., n on each element of b and
denote this as $$ c_i = \sqrt{n} \times (\hat \beta ^ i _{OLS,n} - 2)$$
using OLS. Save it in the memory.

```{r}

n = c(2, 3, 4, 5, 6, 7, 8, 9, 10, 15, 20, 30, 50, 75, 100, 250, 500, 1000, 2000, 3000)

Q4 <- function(n){
  b <- c()
  for (j in c(1:2500)) {

    # a
    x = matrix(runif(n, 0, 12), nrow = n, ncol = 1)
    
    # b
    u = matrix(runif(n, -4, 4), nrow = n, ncol = 1)
    
    # c
    y = 3 + (2 * x) + u
    
    # d
    cov_xy = cov(x, y)
    var_x = var(x)
    beta_hat = cov_xy/var_x
    b = append(b, beta_hat)
  }
  #  f
  var_b = var(b)
  
  c = sqrt(n)*(b - 2)
  return(list(n=n,b=b,var_b=var_b,c=c))
}

for(i in n){
  assign(paste0("q",i),Q4(i))
  hist_data <- hist(Q4(i)$c, breaks = 100,
       main = paste("Histogram of c for n = ", i, sep = ""),
       xlab = "c_i values",
       xlim = c(-3,3))
    x_values <- seq(min(Q4(i)$c), max(Q4(i)$c), length = 100)
    y_values <- dnorm(x_values, mean = mean(Q4(i)$c), sd = sd(Q4(i)$c)) 
    y_values <- y_values * diff(hist_data$mids[1:2]) * length(Q4(i)$c) 
    lines(x_values, y_values, lwd = 2)
}

```

• (h) What happens to the reported values in (f) and the histogram in
(g) as n grows large? Discuss.

5.  (R) Video Game Sales Regression

The dataset for this exercise is available in VideoGamesSales_Main.csv.
This dataset contains information on the global sales and critic and
user review ratings for videogames launched between 2001 and 2012 (from
www.vgchartz.com).

• Your task is to develop a regression model (using lm()) that links
global sales to video game reviews, and explore ways in which the model
fit could be improved through suit- able changes to the model
specification and vdgmriables. (a) (R) First, create a frequency table
of 3 variables: platform, genre, and rating.

```{r}

vdgm <- fread("videogamesales_main.csv",header = T)
if(demo) {str(vdgm)}

ptl_frq <- table(vdgm$Platform)
gnr_frq <- table(vdgm$Genre)
rating_frq <- table(vdgm$Rating)

```

b.  

    (R) Create categorial variables for platform, genre, and rating
        using data.table. Also create a variable for the age of the game
        relative to year 2013 (Note that these games were released
        before 2013).

```{r}

vdgm_2013 <- vdgm[,Platform]
vdgm$Genre<-as.factor(vdgm$Genre)
vdgm$Platform<-as.factor(vdgm$Platform)
vdgm$Rating<-as.factor(vdgm$Rating)

```

(c) 

    (R) Run a regression with all relevant X variables. Report the
        adjusted R-squared.

```{r}

lm1 <- lm(Global_Sales ~. - Name - Publisher - Developer,data = vdgm)
summary(lm1)

```

(d) 

    (R) Now, generate natural log of the following variables: global
        sales, critic_score, critic_count, user_socre, user_count as
        ln\_[original_variable_name].

```{r}

vdgm$ln_Global_Sales = log(vdgm$Global_Sales)
vdgm$ln_Critic_Score = log(vdgm$Critic_Score)
vdgm$ln_Critic_Count = log(vdgm$Critic_Count)
vdgm$ln_User_Score = log(vdgm$User_Score)
vdgm$ln_User_Count = log(vdgm$User_Count)

```

(e) 

    (R) Run a regression with the log of Y variable and report adjusted
        R-squared.

```{r}

lm2 <- lm(ln_Global_Sales ~. - Name - ln_Critic_Score - ln_Critic_Count - ln_User_Score - ln_User_Count - Publisher - Developer - Global_Sales,data = vdgm)
summary(lm2)

```

(f) 

    (R) Run a regression with the log of Y variable as well as log of X
        variables generated in part d). Report adjusted R-squared.

```{r}

lm3 <- lm(ln_Global_Sales ~. - Global_Sales - Name - Publisher - Developer - Critic_Score - Critic_Count -User_Score - ln_User_Count,data = vdgm)
summary(lm3)

```

(g) Which model (out of part c, e, and f) offers the highest adjusted
    R-squared? What would be the economic reasoning on why that
    particular model provides the best fit?

Ans: lm3 has the highest R-squared value. It makes sense that this has the best fit as the sales variable usually been increasing exponentially.

(h) Interpret the parameter estimates for each of 'genre' in plain English. 

Ans: Simulation Games created the highest impact in terms of Global Sales and Sports created the lowest based on the parameter estimates.
    
(i) Interpret the parameter estimate for 'rating' in plain English.

Ans: Simulation Games created the highest impact in terms of Global Sales and Sports created the lowest based on the parameter estimates.

(i) Interpret the parameter estimate for 'ln_user_count' in plain
    English.
    
Ans: ln_User_Count has a p_value < 0.05. There we can reject the null hypothesis i.e. the coefficeint of ln_user_count = 0.